---
title: "Project Modeling"
format: html
editor: visual
---

## Introduction

In this file, we will be continuing the analysis of the diabetes data set from the EDA file. This file will focus on modeling the data using a training data set to fit a classification tree and random forest model. The use of these models will allow us to predict the Diabetes_binary variable. We will finally narrow the selection of models to only our best classification tree and our best random forest model using log-loss with five-fold cross validation, before comparing these final two to declare an overall best model. 

## Data Preparation

Before we can start modeling, we need to split the data into training (70%) and test (30%) sets.
```{r}
library(tidymodels)
set.seed(123)

split <- initial_split(diabetes, prop = 0.7, strata = Diabetes_binary)
train_diabetes <- training(split)
test_diabetes <- testing(split)
```

Next, we will fit a classification tree model to the data using varying values for the parameter and using five-fold CV on the training set to evaluate the best model. We will use the five predictors described in our EDA performed in the previous file. For clarity, a classification tree model splits the feature space into disjoint regions to create a tree of decisions, with each node having a predicted class distribution. The class with the highest probability is the predicted label.
```{r}

```
